---
title: "visualization"
format: html
editor: visual
---

```{r}
library(ggplot2)
library(tidyverse)
library(ggsci)
library(ggstar)
library(ggpubr)
library(RColorBrewer)
library(cowplot)
library(reshape2)
```

```{r}
library(GEOquery)
```



```{rheader```{r}
df <- read.csv('pca_combat.csv')
meta <- read.delim("a_bulkRNA_Kundaje.txt")
df <- inner_join(df, meta, by = join_by(Var4 == Sample.Name))
p <- df %>% 
  filter(str_detect(Var3, 'APS|MPS|H7hESC|H7Trzl|DLL1pPXM|LtM')) %>% 
  ggplot(aes(x = pc1, y = pc2)) + 
  geom_star(aes(starshape = Var3, fill = Var4, color = factor(Comment.batch.)), size = 3, starstroke = 0.7) +
  theme_pubr() +
  labs(
    x = 'PC1',
    y = 'PC2',
    starshape = 'cell type',
    fill = 'sample',
    color = 'batch'
  ) +
  guides(fill = guide_legend(ncol=3), starshape = guide_legend(ncol=3), color = guide_legend(ncol=3)) + 
  theme(
    text = element_text(size = 20, family = 'arial', face = 'bold'),
    legend.position = "right",
    legend.text = element_text(size = 12, family = 'arial', face = 'bold'),
    legend.title = element_text(size = 12, family = 'arial', face = 'bold'),
    # legend.position.inside = c(0.35, 0.5),
    aspect.ratio = 1,
    axis.ticks.length = unit(-1.25, "mm"),
    axis.ticks = element_line(size = 0.75),
    axis.line = element_line(size = 0.75, lineend = 'round'),
    legend.background = element_blank(),
    legend.box.background = element_rect(colour = "black", size = 0.75),
    legend.key.spacing.y = unit(2, 'mm'),
    legend.key.width = unit(2, 'mm'),
    panel.border = element_blank(),
  ) +
  scale_color_lancet()
save_plot('pca_after_correction.png', p, base_width = 8, base_height = 5, dpi = 300)
```

```{r}
df <- read.csv('pca.csv')
meta <- read.delim("a_bulkRNA_Kundaje.txt")
df <- inner_join(df, meta, by = join_by(Var4 == Sample.Name))
p <- df %>% 
  filter(str_detect(Var3, 'APS|MPS|H7hESC|H7Trzl|DLL1pPXM|LtM')) %>% 
  ggplot(aes(x = pc1, y = pc2)) + 
  geom_star(aes(starshape = Var3, fill = Var4, color = factor(Comment.batch.)), size = 3, starstroke = 0.7) +
  theme_pubr() +
  labs(
    x = 'PC1',
    y = 'PC2',
    starshape = 'cell type',
    fill = 'sample',
    color = 'batch'
  ) +
  guides(fill = guide_legend(ncol=3), starshape = guide_legend(ncol=3), color = guide_legend(ncol=3)) + 
  theme(
    text = element_text(size = 20, family = 'arial', face = 'bold'),
    legend.position = "right",
    legend.text = element_text(size = 12, family = 'arial', face = 'bold'),
    legend.title = element_text(size = 12, family = 'arial', face = 'bold'),
    # legend.position.inside = c(0.35, 0.5),
    aspect.ratio = 1,
    axis.ticks.length = unit(-1.25, "mm"),
    axis.ticks = element_line(size = 0.75),
    axis.line = element_line(size = 0.75, lineend = 'round'),
    legend.background = element_blank(),
    legend.box.background = element_rect(colour = "black", size = 0.75),
    legend.key.spacing.y = unit(2, 'mm'),
    legend.key.width = unit(2, 'mm'),
    panel.border = element_blank(),
  ) +
  scale_color_lancet()
save_plot('pca_before_correction.png', p, base_width = 8, base_height = 5, dpi = 300)
```

```{r}
loh_RNA <- read.delim("./GSM2257301_All_samples_bulk_tpm.txt/GSM2257301_All_samples_bulk_tpm.txt")
# loh_RNA %>% write_csv('loh_RNA.csv')
loh_RNA <- loh_RNA %>%
  filter(
    str_detect(geneSymbol, 'EOMES|FOXF1|NKX2-5|ISL1|TBX6|HAND1|FOXC2|TBX1|TBX5|MSGN1|APLNR')
  ) %>% 
  dplyr::select(
    matches('^APS|^MPS|hESC|pPXM|LtM|geneSym|Trzl|GARPpCrdcM|D3EarlySmt')
  ) %>% 
  melt(id.vars = 'geneSymbol') %>% 
  rename(gene = geneSymbol, sample = variable, tpm = value) %>% 
  mutate(log21p_tpm = log2(tpm + 1)) %>% 
  {left_join(., meta, by = join_by(sample == Sample.Name))} %>% 
  select(gene, sample, tpm, log21p_tpm, Comment.batch.) %>% 
  rename(batch = Comment.batch.) %>% 
  mutate(batch = factor(batch, levels = c(1, 2, 3), labels = c('batch 1', 'batch 2', 'batch 3'))) %>% 
  mutate(sample = gsub('DLLp', 'DLL1p', sample)) %>% 
  mutate(sample = gsub('APS\\_\\d', 'APS', sample)) %>% 
  mutate(sample = gsub('DLL1pPXM\\_\\d', 'DLL1PXM', sample)) %>% 
  mutate(sample = gsub('H7hESC\\_\\d', 'H7hESC', sample)) %>% 
  mutate(sample = gsub('MPS\\_\\d', 'MPS', sample)) %>%
  mutate(sample = gsub('D2LtM\\_\\d', 'D2LtM', sample)) %>% 
  mutate(sample = gsub('H7hESC|H7Trzl', 'hESC', sample)) %>%
  mutate(sample = gsub('D3GARPpCrdcM\\_\\d', 'D3GARPpCrdcM', sample)) %>%
  mutate(sample = gsub('D3EarlySmt\\_\\d', 'D3EarlySmt', sample)) %>%
  mutate(batch = factor(ifelse(sample == 'D3GARPpCrdcM', 3, batch), levels = c(1, 2, 3), labels = c('batch 1', 'batch 2', 'batch 3'))) %>% 
  group_by(gene, sample, batch) %>% 
  mutate(mean_tpm = mean(tpm), mean_log21p_tpm = mean(log21p_tpm)) #%>% mutate(mean_tpm = ifelse(sample == 'APS', NA, mean_tpm), mean_log21p_tpm = ifelse(sample == 'APS', NA, mean_log21p_tpm))
```

```{r}
library(dplyr)
library(stringr)
library(ggplot2)
library(ggstar)
library(ggpubr)
library(patchwork)
library(grid)

loh_RNA_lpm <- loh_RNA %>%
  filter(str_detect(sample, "hESC|MPS|D2Lt|CrdcM")) %>%
  mutate(sample = ordered(sample, levels = c("hESC", "MPS", "D2LtM", "D3GARPpCrdcM")))

# slots for custom x tick labels
xlab_p1 <- c(
  "hESC" = "hESC (D0)",
  "MPS" = "MPS (D1)",
  "D2LtM" = "Lateral meso (D2)",
  "D3GARPpCrdcM" = "Cardiac meso (D3)"
)

xlab_p2 <- c(
  "hESC" = "hESC (D0)",
  "APS" = "APS (D1)",
  "DLL1PXM" = "Paraxial meso (D2)",
  "D3EarlySmt" = "Early somites (D3)"
)

base_theme <- theme_pubr() +
  theme(
    text = element_text(size = 20, family = "arial", face = "bold"),
    legend.position = "none",
    legend.text = element_text(size = 12, family = "arial", face = "bold"),
    legend.title = element_text(size = 12, family = "arial", face = "bold"),
    legend.position.inside = c(0.85, 0.5),
    aspect.ratio = 1,
    axis.ticks.length = unit(-1.25, "mm"),
    axis.ticks = element_line(linewidth = 0.75),
    axis.line = element_line(linewidth = 0.75, lineend = "round"),
    legend.background = element_blank(),
    legend.box.background = element_rect(colour = "black", linewidth = 0.75),
    legend.key.spacing.y = unit(1, "mm"),
    legend.key.width = unit(2, "mm"),
    axis.text.x = element_text(angle = 30, vjust = 0.8, hjust = 0.91),
    panel.border = element_blank()
  )

p1 <- loh_RNA %>%
  filter(str_detect(sample, "hESC|MPS|D2Lt|CrdcM")) %>%
  mutate(sample = ordered(sample, levels = c("hESC", "MPS", "D2LtM", "D3GARPpCrdcM"))) %>%
  filter(gene == "APLNR", batch == "batch 3") %>%
  ggplot(aes(x = sample, y = log21p_tpm)) +
  geom_star(
    starshape = 15, fill = "#00bbd7", color = "#077fb0",
    size = 3, starstroke = 1, alpha = 0.75
  ) +
  geom_line(
    aes(group = interaction(batch, gene), y = mean_log21p_tpm),
    color = "#00468BFF", alpha = 0.75, linewidth = 1,
    data = filter(loh_RNA_lpm, gene == "APLNR", batch == "batch 3")
  ) +
  scale_x_discrete(labels = xlab_p1) +
  coord_cartesian(ylim = c(0, 10)) +
  labs(
    y = expression(bold(log["2"] * ("TPM + 1"))),
    x = NULL,
    title = "APLNR"
  ) +
  base_theme

p2 <- loh_RNA %>%
  filter(str_detect(sample, "hESC|APS|DLL1PXM|D3Early")) %>%
  mutate(sample = ordered(sample, levels = c("hESC", "APS", "DLL1PXM", "D3EarlySmt"))) %>%
  filter(gene == "APLNR", str_detect(batch, "batch 2|batch 1")) %>%
  ggplot(aes(x = sample, y = log21p_tpm)) +
  geom_star(
    starshape = 15, fill = "#00bbd7", color = "#077fb0",
    size = 3, starstroke = 1, alpha = 0.75
  ) +
  geom_line(
    aes(group = gene, y = mean_log21p_tpm),
    color = "#00468BFF", alpha = 0.75, linewidth = 1
  ) +
  scale_x_discrete(labels = xlab_p2) +
  coord_cartesian(ylim = c(0, 10)) +
  labs(
    y = expression(bold(log["2"] * ("TPM + 1"))),
    x = NULL,
    title = "APLNR"
  ) +
  base_theme

p <- p1 + p2 + plot_layout(ncol = 2)
save_plot('APLNR_loh.png', p, base_width = 10, base_height = 5, dpi = 300)
```

```{r}
p <- loh_RNA %>%
  filter(str_detect(sample, 'hESC|MPS|D2Lt|APS|CrdcM')) %>% 
  mutate(sample = ordered(sample, levels = c('hESC', 'MPS', 'D2LtM', 'D3GARPpCrdcM', 'APS'))) %>% 
  filter(gene == 'EOMES', batch == 'batch 3') %>% 
  ggplot(aes(x = sample, y = tpm)) +
  geom_star(starshape = 15, fill = '#00bbd7', color = '#077fb0', size = 3, starstroke = 1, alpha = 0.75) +
  geom_line(aes(group = interaction(batch, gene), color = gene, y = mean_tpm), linetype = 'dashed') +
  labs(
    title = 'EOMES'
  ) +
  theme_pubr() +
  theme(
    text = element_text(size = 20, family = 'arial', face = 'bold'),
    legend.position = "none",
    legend.text = element_text(size = 12, family = 'arial', face = 'bold'),
    legend.title = element_text(size = 12, family = 'arial', face = 'bold'),
    legend.position.inside = c(0.85, 0.5),
    aspect.ratio = 1,
    axis.ticks.length = unit(-1.25, "mm"),
    axis.ticks = element_line(size = 0.75),
    axis.line = element_line(size = 0.75, lineend = 'round'),
    legend.background = element_blank(),
    legend.box.background = element_rect(colour = "black", size = 0.75),
    legend.key.spacing.y = unit(1, 'mm'),
    legend.key.width = unit(2, 'mm'),
    panel.border = element_blank(),
  )
save_plot('hESC_MPS_LtM_EOMES.png', p, base_width = 5, base_height = 5, dpi = 300)

p <- loh_RNA %>%
  filter(str_detect(sample, 'hESC|MPS|D2Lt|APS|CrdcM')) %>% 
  mutate(sample = ordered(sample, levels = c('hESC', 'MPS', 'D2LtM', 'D3GARPpCrdcM', 'APS'))) %>% 
  filter(gene == 'FOXF1', batch == 'batch 3') %>% 
  ggplot(aes(x = sample, y = tpm)) +
  geom_star(starshape = 15, fill = '#00bbd7', color = '#077fb0', size = 3, starstroke = 1, alpha = 0.75) +
  geom_line(aes(group = interaction(batch, gene), color = gene, y = mean_tpm), linetype = 'dashed') +
  labs(
    title = 'FOXF1'
  ) +
  theme_pubr() +
  theme(
    text = element_text(size = 20, family = 'arial', face = 'bold'),
    legend.position = "none",
    legend.text = element_text(size = 12, family = 'arial', face = 'bold'),
    legend.title = element_text(size = 12, family = 'arial', face = 'bold'),
    legend.position.inside = c(0.85, 0.5),
    aspect.ratio = 1,
    axis.ticks.length = unit(-1.25, "mm"),
    axis.ticks = element_line(size = 0.75),
    axis.line = element_line(size = 0.75, lineend = 'round'),
    legend.background = element_blank(),
    legend.box.background = element_rect(colour = "black", size = 0.75),
    legend.key.spacing.y = unit(1, 'mm'),
    legend.key.width = unit(2, 'mm'),
    panel.border = element_blank(),
  )
save_plot('hESC_MPS_LtM_FOXF1.png', p, base_width = 5, base_height = 5, dpi = 300)

p <- loh_RNA %>%
  filter(str_detect(sample, 'hESC|MPS|D2Lt|APS|CrdcM')) %>% 
  mutate(sample = ordered(sample, levels = c('hESC', 'MPS', 'D2LtM', 'D3GARPpCrdcM', 'APS'))) %>% 
  filter(gene == 'ISL1', batch == 'batch 3') %>% 
  ggplot(aes(x = sample, y = tpm)) +
  geom_star(starshape = 15, fill = '#00bbd7', color = '#077fb0', size = 3, starstroke = 1, alpha = 0.75) +
  geom_line(aes(group = interaction(batch, gene), color = gene, y = mean_tpm), linetype = 'dashed') +
  labs(
    title = 'IS1'
  ) +
  theme_pubr() +
  theme(
    text = element_text(size = 20, family = 'arial', face = 'bold'),
    legend.position = "none",
    legend.text = element_text(size = 12, family = 'arial', face = 'bold'),
    legend.title = element_text(size = 12, family = 'arial', face = 'bold'),
    legend.position.inside = c(0.85, 0.5),
    aspect.ratio = 1,
    axis.ticks.length = unit(-1.25, "mm"),
    axis.ticks = element_line(size = 0.75),
    axis.line = element_line(size = 0.75, lineend = 'round'),
    legend.background = element_blank(),
    legend.box.background = element_rect(colour = "black", size = 0.75),
    legend.key.spacing.y = unit(1, 'mm'),
    legend.key.width = unit(2, 'mm'),
    panel.border = element_blank(),
  )
save_plot('hESC_MPS_LtM_ISL1.png', p, base_width = 5, base_height = 5, dpi = 300)

p <- loh_RNA %>%
  filter(str_detect(sample, 'hESC|MPS|D2Lt|APS|CrdcM')) %>% 
  mutate(sample = ordered(sample, levels = c('hESC', 'MPS', 'D2LtM', 'D3GARPpCrdcM', 'APS'))) %>% 
  filter(gene == 'NKX2-5', batch == 'batch 3') %>% 
  ggplot(aes(x = sample, y = tpm)) +
  geom_star(starshape = 15, fill = '#00bbd7', color = '#077fb0', size = 3, starstroke = 1, alpha = 0.75) +
  geom_line(aes(group = interaction(batch, gene), color = gene, y = mean_tpm), linetype = 'dashed') +
  labs(
    title = 'NKX2-5'
  ) +
  theme_pubr() +
  theme(
    text = element_text(size = 20, family = 'arial', face = 'bold'),
    legend.position = "none",
    legend.text = element_text(size = 12, family = 'arial', face = 'bold'),
    legend.title = element_text(size = 12, family = 'arial', face = 'bold'),
    legend.position.inside = c(0.85, 0.5),
    aspect.ratio = 1,
    axis.ticks.length = unit(-1.25, "mm"),
    axis.ticks = element_line(size = 0.75),
    axis.line = element_line(size = 0.75, lineend = 'round'),
    legend.background = element_blank(),
    legend.box.background = element_rect(colour = "black", size = 0.75),
    legend.key.spacing.y = unit(1, 'mm'),
    legend.key.width = unit(2, 'mm'),
    panel.border = element_blank(),
  )
save_plot('hESC_MPS_LtM_NKX2-5.png', p, base_width = 5, base_height = 5, dpi = 300)

p <- loh_RNA %>%
  filter(str_detect(sample, 'hESC|MPS|D2Lt|APS|CrdcM')) %>% 
  mutate(sample = ordered(sample, levels = c('hESC', 'MPS', 'D2LtM', 'D3GARPpCrdcM', 'APS'))) %>% 
  filter(gene == 'TBX1', batch == 'batch 3') %>% 
  ggplot(aes(x = sample, y = tpm)) +
  geom_star(starshape = 15, fill = '#00bbd7', color = '#077fb0', size = 3, starstroke = 1, alpha = 0.75) +
  geom_line(aes(group = interaction(batch, gene), color = gene, y = mean_tpm), linetype = 'dashed') +
  labs(
    title = 'TBX1'
  ) +
  theme_pubr() +
  theme(
    text = element_text(size = 20, family = 'arial', face = 'bold'),
    legend.position = "none",
    legend.text = element_text(size = 12, family = 'arial', face = 'bold'),
    legend.title = element_text(size = 12, family = 'arial', face = 'bold'),
    legend.position.inside = c(0.85, 0.5),
    aspect.ratio = 1,
    axis.ticks.length = unit(-1.25, "mm"),
    axis.ticks = element_line(size = 0.75),
    axis.line = element_line(size = 0.75, lineend = 'round'),
    legend.background = element_blank(),
    legend.box.background = element_rect(colour = "black", size = 0.75),
    legend.key.spacing.y = unit(1, 'mm'),
    legend.key.width = unit(2, 'mm'),
    panel.border = element_blank(),
  )
save_plot('hESC_MPS_LtM_TBX1.png', p, base_width = 5, base_height = 5, dpi = 300)

p <- loh_RNA %>%
  filter(str_detect(sample, 'hESC|MPS|D2Lt|APS|CrdcM')) %>% 
  mutate(sample = ordered(sample, levels = c('hESC', 'MPS', 'D2LtM', 'D3GARPpCrdcM', 'APS'))) %>% 
  filter(gene == 'TBX5', batch == 'batch 3') %>% 
  ggplot(aes(x = sample, y = tpm)) +
  geom_star(starshape = 15, fill = '#00bbd7', color = '#077fb0', size = 3, starstroke = 1, alpha = 0.75) +
  geom_line(aes(group = interaction(batch, gene), color = gene, y = mean_tpm), linetype = 'dashed') +
  labs(
    title = 'TBX5'
  ) +
  theme_pubr() +
  theme(
    text = element_text(size = 20, family = 'arial', face = 'bold'),
    legend.position = "none",
    legend.text = element_text(size = 12, family = 'arial', face = 'bold'),
    legend.title = element_text(size = 12, family = 'arial', face = 'bold'),
    legend.position.inside = c(0.85, 0.5),
    aspect.ratio = 1,
    axis.ticks.length = unit(-1.25, "mm"),
    axis.ticks = element_line(size = 0.75),
    axis.line = element_line(size = 0.75, lineend = 'round'),
    legend.background = element_blank(),
    legend.box.background = element_rect(colour = "black", size = 0.75),
    legend.key.spacing.y = unit(1, 'mm'),
    legend.key.width = unit(2, 'mm'),
    panel.border = element_blank(),
  )
save_plot('hESC_MPS_LtM_TBX5.png', p, base_width = 5, base_height = 5, dpi = 300)
```


```{r}
loh_RNA_lpm <- loh_RNA %>%
  filter(str_detect(sample, 'hESC|MPS|D2Lt|CrdcM')) %>% 
  mutate(sample = ordered(sample, levels = c('hESC', 'MPS', 'D2LtM', 'D3GARPpCrdcM'))) 

p <- loh_RNA %>%
  filter(str_detect(sample, 'hESC|MPS|D2Lt|APS|CrdcM')) %>% 
  mutate(sample = ordered(sample, levels = c('hESC', 'MPS', 'D2LtM', 'D3GARPpCrdcM', 'APS'))) %>% 
  filter(gene == 'EOMES', batch == 'batch 3') %>% 
  ggplot(aes(x = sample, y = log21p_tpm)) +
  geom_star(starshape = 15, fill = '#00bbd7', color = '#077fb0', size = 3, starstroke = 1, alpha = 0.75) +
  geom_line(aes(group = interaction(batch, gene), y = mean_log21p_tpm), color = '#00468BFF', alpha = 0.75, linewidth = 1, data = filter(loh_RNA_lpm, gene == 'EOMES', batch == 'batch 3')) +
  labs(
    y=expression(bold(log["2"] * ("TPM + 1"))),
    title = 'EOMES'
  ) +
  theme_pubr() +
  theme(
    text = element_text(size = 20, family = 'arial', face = 'bold'),
    legend.position = "none",
    legend.text = element_text(size = 12, family = 'arial', face = 'bold'),
    legend.title = element_text(size = 12, family = 'arial', face = 'bold'),
    legend.position.inside = c(0.85, 0.5),
    aspect.ratio = 1,
    axis.ticks.length = unit(-1.25, "mm"),
    axis.ticks = element_line(size = 0.75),
    axis.line = element_line(size = 0.75, lineend = 'round'),
    legend.background = element_blank(),
    legend.box.background = element_rect(colour = "black", size = 0.75),
    legend.key.spacing.y = unit(1, 'mm'),
    legend.key.width = unit(2, 'mm'),
    axis.text.x = element_text(angle = 30, vjust = 0.8, hjust=0.91),
    panel.border = element_blank(),
  )
save_plot('hESC_MPS_LtM_EOMES_log2_1p.png', p, base_width = 5, base_height = 5, dpi = 300)

p <- loh_RNA %>%
  filter(str_detect(sample, 'hESC|MPS|D2Lt|APS|CrdcM')) %>% 
  mutate(sample = ordered(sample, levels = c('hESC', 'MPS', 'D2LtM', 'D3GARPpCrdcM', 'APS'))) %>% 
  filter(gene == 'FOXF1', batch == 'batch 3') %>% 
  ggplot(aes(x = sample, y = log21p_tpm)) +
  geom_star(starshape = 15, fill = '#00bbd7', color = '#077fb0', size = 3, starstroke = 1, alpha = 0.75) +
  geom_line(aes(group = interaction(batch, gene), y = mean_log21p_tpm), color = '#00468BFF', alpha = 0.75, linewidth = 1, data = filter(loh_RNA_lpm, gene == 'FOXF1', batch == 'batch 3')) +
  labs(
    y=expression(bold(log["2"] * ("TPM + 1"))),
    title = 'FOXF1'
  ) +
  theme_pubr() +
  theme(
    text = element_text(size = 20, family = 'arial', face = 'bold'),
    legend.position = "none",
    legend.text = element_text(size = 12, family = 'arial', face = 'bold'),
    legend.title = element_text(size = 12, family = 'arial', face = 'bold'),
    legend.position.inside = c(0.85, 0.5),
    aspect.ratio = 1,
    axis.ticks.length = unit(-1.25, "mm"),
    axis.ticks = element_line(size = 0.75),
    axis.line = element_line(size = 0.75, lineend = 'round'),
    legend.background = element_blank(),
    legend.box.background = element_rect(colour = "black", size = 0.75),
    legend.key.spacing.y = unit(1, 'mm'),
    legend.key.width = unit(2, 'mm'),
    axis.text.x = element_text(angle = 30, vjust = 0.8, hjust=0.91),
    panel.border = element_blank(),
  )
save_plot('hESC_MPS_LtM_FOXF1_log2_1p.png', p, base_width = 5, base_height = 5, dpi = 300)

p <- loh_RNA %>%
  filter(str_detect(sample, 'hESC|MPS|D2Lt|APS|CrdcM')) %>% 
  mutate(sample = ordered(sample, levels = c('hESC', 'MPS', 'D2LtM', 'D3GARPpCrdcM', 'APS'))) %>% 
  filter(gene == 'ISL1', batch == 'batch 3') %>% 
  ggplot(aes(x = sample, y = log21p_tpm)) +
  geom_star(starshape = 15, fill = '#00bbd7', color = '#077fb0', size = 3, starstroke = 1, alpha = 0.75) +
  geom_line(aes(group = interaction(batch, gene), y = mean_log21p_tpm), color = '#00468BFF', alpha = 0.75, linewidth = 1, data = filter(loh_RNA_lpm, gene == 'ISL1', batch == 'batch 3')) +
  labs(
    y=expression(bold(log["2"] * ("TPM + 1"))),
    title = 'ISL1'
  ) +
  theme_pubr() +
  theme(
    text = element_text(size = 20, family = 'arial', face = 'bold'),
    legend.position = "none",
    legend.text = element_text(size = 12, family = 'arial', face = 'bold'),
    legend.title = element_text(size = 12, family = 'arial', face = 'bold'),
    legend.position.inside = c(0.85, 0.5),
    aspect.ratio = 1,
    axis.ticks.length = unit(-1.25, "mm"),
    axis.ticks = element_line(size = 0.75),
    axis.line = element_line(size = 0.75, lineend = 'round'),
    legend.background = element_blank(),
    legend.box.background = element_rect(colour = "black", size = 0.75),
    legend.key.spacing.y = unit(1, 'mm'),
    legend.key.width = unit(2, 'mm'),
    axis.text.x = element_text(angle = 30, vjust = 0.8, hjust=0.91),
    panel.border = element_blank(),
  )
save_plot('hESC_MPS_LtM_ISL1_log2_1p.png', p, base_width = 5, base_height = 5, dpi = 300)

p <- loh_RNA %>%
  filter(str_detect(sample, 'hESC|MPS|D2Lt|APS|CrdcM')) %>% 
  mutate(sample = ordered(sample, levels = c('hESC', 'MPS', 'D2LtM', 'D3GARPpCrdcM', 'APS'))) %>% 
  filter(gene == 'NKX2-5', batch == 'batch 3') %>% 
  ggplot(aes(x = sample, y = log21p_tpm)) +
  geom_star(starshape = 15, fill = '#00bbd7', color = '#077fb0', size = 3, starstroke = 1, alpha = 0.75) +
  geom_line(aes(group = interaction(batch, gene), y = mean_log21p_tpm), color = '#00468BFF', alpha = 0.75, linewidth = 1, data = filter(loh_RNA_lpm, gene == 'NKX2-5', batch == 'batch 3')) +
  labs(
    y=expression(bold(log["2"] * ("TPM + 1"))),
    title = 'NKX2-5'
  ) +
  theme_pubr() +
  scale_y_continuous(breaks = c(0, 1, 2)) +
  theme(
    text = element_text(size = 20, family = 'arial', face = 'bold'),
    legend.position = "none",
    legend.text = element_text(size = 12, family = 'arial', face = 'bold'),
    legend.title = element_text(size = 12, family = 'arial', face = 'bold'),
    legend.position.inside = c(0.85, 0.5),
    aspect.ratio = 1,
    axis.ticks.length = unit(-1.25, "mm"),
    axis.ticks = element_line(size = 0.75),
    axis.line = element_line(size = 0.75, lineend = 'round'),
    legend.background = element_blank(),
    legend.box.background = element_rect(colour = "black", size = 0.75),
    legend.key.spacing.y = unit(1, 'mm'),
    legend.key.width = unit(2, 'mm'),
    axis.text.x = element_text(angle = 30, vjust = 0.8, hjust=0.91),
    panel.border = element_blank(),
  )
save_plot('hESC_MPS_LtM_NKX2-5_log2_1p.png', p, base_width = 5, base_height = 5, dpi = 300)

p <- loh_RNA %>%
  filter(str_detect(sample, 'hESC|MPS|D2Lt|APS|CrdcM')) %>% 
  mutate(sample = ordered(sample, levels = c('hESC', 'MPS', 'D2LtM', 'D3GARPpCrdcM', 'APS'))) %>% 
  filter(gene == 'TBX1', batch == 'batch 3') %>% 
  ggplot(aes(x = sample, y = log21p_tpm)) +
  geom_star(starshape = 15, fill = '#00bbd7', color = '#077fb0', size = 3, starstroke = 1, alpha = 0.75) +
  geom_line(aes(group = interaction(batch, gene), y = mean_log21p_tpm), color = '#00468BFF', alpha = 0.75, linewidth = 1, data = filter(loh_RNA_lpm, gene == 'TBX1', batch == 'batch 3')) +
  labs(
    y=expression(bold(log["2"] * ("TPM + 1"))),
    title = 'TBX1'
  ) +
  theme_pubr() +
  theme(
    text = element_text(size = 20, family = 'arial', face = 'bold'),
    legend.position = "none",
    legend.text = element_text(size = 12, family = 'arial', face = 'bold'),
    legend.title = element_text(size = 12, family = 'arial', face = 'bold'),
    legend.position.inside = c(0.85, 0.5),
    aspect.ratio = 1,
    axis.ticks.length = unit(-1.25, "mm"),
    axis.ticks = element_line(size = 0.75),
    axis.line = element_line(size = 0.75, lineend = 'round'),
    legend.background = element_blank(),
    legend.box.background = element_rect(colour = "black", size = 0.75),
    legend.key.spacing.y = unit(1, 'mm'),
    legend.key.width = unit(2, 'mm'),
    axis.text.x = element_text(angle = 30, vjust = 0.8, hjust=0.91),
    panel.border = element_blank(),
  )
save_plot('hESC_MPS_LtM_TBX1_log2_1p.png', p, base_width = 5, base_height = 5, dpi = 300)

p <- loh_RNA %>%
  filter(str_detect(sample, 'hESC|MPS|D2Lt|APS|CrdcM')) %>% 
  mutate(sample = ordered(sample, levels = c('hESC', 'MPS', 'D2LtM', 'D3GARPpCrdcM', 'APS'))) %>% 
  filter(gene == 'TBX5', batch == 'batch 3') %>% 
  ggplot(aes(x = sample, y = log21p_tpm)) +
  geom_star(starshape = 15, fill = '#00bbd7', color = '#077fb0', size = 3, starstroke = 1, alpha = 0.75) +
  geom_line(aes(group = interaction(batch, gene), y = mean_log21p_tpm), color = '#00468BFF', alpha = 0.75, linewidth = 1, data = filter(loh_RNA_lpm, gene == 'TBX5', batch == 'batch 3')) +
  labs(
    y=expression(bold(log["2"] * ("TPM + 1"))),
    title = 'TBX5'
  ) +
  theme_pubr() +
  theme(
    text = element_text(size = 20, family = 'arial', face = 'bold'),
    legend.position = "none",
    legend.text = element_text(size = 12, family = 'arial', face = 'bold'),
    legend.title = element_text(size = 12, family = 'arial', face = 'bold'),
    legend.position.inside = c(0.85, 0.5),
    aspect.ratio = 1,
    axis.ticks.length = unit(-1.25, "mm"),
    axis.ticks = element_line(size = 0.75),
    axis.line = element_line(size = 0.75, lineend = 'round'),
    legend.background = element_blank(),
    legend.box.background = element_rect(colour = "black", size = 0.75),
    legend.key.spacing.y = unit(1, 'mm'),
    legend.key.width = unit(2, 'mm'),
    axis.text.x = element_text(angle = 30, vjust = 0.8, hjust=0.91),
    panel.border = element_blank(),
  )
save_plot('hESC_MPS_LtM_TBX5_log2_1p.png', p, base_width = 5, base_height = 5, dpi = 300)

p <- loh_RNA %>%
  filter(str_detect(sample, 'hESC|MPS|D2Lt|APS|CrdcM')) %>% 
  mutate(sample = ordered(sample, levels = c('hESC', 'MPS', 'D2LtM', 'D3GARPpCrdcM', 'APS'))) %>% 
  filter(gene == 'TBX6', batch == 'batch 3') %>% 
  ggplot(aes(x = sample, y = log21p_tpm)) +
  geom_star(starshape = 15, fill = '#00bbd7', color = '#077fb0', size = 3, starstroke = 1, alpha = 0.75) +
  geom_line(aes(group = interaction(batch, gene), y = mean_log21p_tpm), color = '#00468BFF', alpha = 0.75, linewidth = 1, data = filter(loh_RNA_lpm, gene == 'TBX6', batch == 'batch 3')) +
  labs(
    y=expression(bold(log["2"] * ("TPM + 1"))),
    title = 'TBX6'
  ) +
  theme_pubr() +
  theme(
    text = element_text(size = 20, family = 'arial', face = 'bold'),
    legend.position = "none",
    legend.text = element_text(size = 12, family = 'arial', face = 'bold'),
    legend.title = element_text(size = 12, family = 'arial', face = 'bold'),
    legend.position.inside = c(0.85, 0.5),
    aspect.ratio = 1,
    axis.ticks.length = unit(-1.25, "mm"),
    axis.ticks = element_line(size = 0.75),
    axis.line = element_line(size = 0.75, lineend = 'round'),
    legend.background = element_blank(),
    legend.box.background = element_rect(colour = "black", size = 0.75),
    legend.key.spacing.y = unit(1, 'mm'),
    legend.key.width = unit(2, 'mm'),
    axis.text.x = element_text(angle = 30, vjust = 0.8, hjust=0.91),
    panel.border = element_blank(),
  )
save_plot('hESC_MPS_LtM_TBX6_log2_1p.png', p, base_width = 5, base_height = 5, dpi = 300)

p <- loh_RNA %>%
  filter(str_detect(sample, 'hESC|MPS|D2Lt|APS|CrdcM')) %>% 
  mutate(sample = ordered(sample, levels = c('hESC', 'MPS', 'D2LtM', 'D3GARPpCrdcM', 'APS'))) %>% 
  filter(gene == 'MSGN1', batch == 'batch 3') %>% 
  ggplot(aes(x = sample, y = log21p_tpm)) +
  geom_star(starshape = 15, fill = '#00bbd7', color = '#077fb0', size = 3, starstroke = 1, alpha = 0.75) +
  geom_line(aes(group = interaction(batch, gene), y = mean_log21p_tpm), color = '#00468BFF', alpha = 0.75, linewidth = 1, data = filter(loh_RNA_lpm, gene == 'MSGN1', batch == 'batch 3')) +
  labs(
    y=expression(bold(log["2"] * ("TPM + 1"))),
    title = 'MSGN1'
  ) +
  theme_pubr() +
  theme(
    text = element_text(size = 20, family = 'arial', face = 'bold'),
    legend.position = "none",
    legend.text = element_text(size = 12, family = 'arial', face = 'bold'),
    legend.title = element_text(size = 12, family = 'arial', face = 'bold'),
    legend.position.inside = c(0.85, 0.5),
    aspect.ratio = 1,
    axis.ticks.length = unit(-1.25, "mm"),
    axis.ticks = element_line(size = 0.75),
    axis.line = element_line(size = 0.75, lineend = 'round'),
    legend.background = element_blank(),
    legend.box.background = element_rect(colour = "black", size = 0.75),
    legend.key.spacing.y = unit(1, 'mm'),
    legend.key.width = unit(2, 'mm'),
    axis.text.x = element_text(angle = 30, vjust = 0.8, hjust=0.91),
    panel.border = element_blank(),
  )
save_plot('hESC_MPS_LtM_MSGN1_log2_1p.png', p, base_width = 5, base_height = 5, dpi = 300)
```


```{r}
p <- loh_RNA %>%
  filter(str_detect(sample, 'hESC|APS|DLL1PXM|D3Early')) %>% 
  mutate(sample = ordered(sample, levels = c('hESC', 'APS', 'DLL1PXM', 'D3EarlySmt'))) %>% 
  filter(gene == 'TBX6', str_detect(batch, 'batch 2|batch 1')) %>% 
  ggplot(aes(x = sample, y = log21p_tpm)) +
  geom_star(starshape = 15, fill = '#00bbd7', color = '#077fb0', size = 3, starstroke = 1, alpha = 0.75) +
  geom_line(aes(group = gene, y = mean_log21p_tpm), color = '#00468BFF', alpha = 0.75, linewidth = 1) +
  labs(
    y=expression(bold(log["2"] * ("TPM + 1"))),
    title = 'TBX6'
  ) +
  theme_pubr() +
  theme(
    text = element_text(size = 20, family = 'arial', face = 'bold'),
    legend.position = "none",
    legend.text = element_text(size = 12, family = 'arial', face = 'bold'),
    legend.title = element_text(size = 12, family = 'arial', face = 'bold'),
    legend.position.inside = c(0.85, 0.5),
    aspect.ratio = 1,
    axis.ticks.length = unit(-1.25, "mm"),
    axis.ticks = element_line(size = 0.75),
    axis.line = element_line(size = 0.75, lineend = 'round'),
    legend.background = element_blank(),
    legend.box.background = element_rect(colour = "black", size = 0.75),
    legend.key.spacing.y = unit(1, 'mm'),
    legend.key.width = unit(2, 'mm'),
    axis.text.x = element_text(angle = 30, vjust = 0.8, hjust=0.91),
    panel.border = element_blank(),
  )
save_plot('hESC_APS_PXM_SMT_TBX6_log2_1p.png', p, base_width = 5, base_height = 5, dpi = 300)

p <- loh_RNA %>%
  filter(str_detect(sample, 'hESC|APS|DLL1PXM|D3Early')) %>% 
  mutate(sample = ordered(sample, levels = c('hESC', 'APS', 'DLL1PXM', 'D3EarlySmt'))) %>% 
  filter(gene == 'HAND1', str_detect(batch, 'batch 2|batch 1')) %>% 
  ggplot(aes(x = sample, y = log21p_tpm)) +
  geom_star(starshape = 15, fill = '#00bbd7', color = '#077fb0', size = 3, starstroke = 1, alpha = 0.75) +
  geom_line(aes(group = gene, y = mean_log21p_tpm), color = '#00468BFF', alpha = 0.75, linewidth = 1) +
  labs(
    y=expression(bold(log["2"] * ("TPM + 1"))),
    title = 'HAND1'
  ) +
  theme_pubr() +
  theme(
    text = element_text(size = 20, family = 'arial', face = 'bold'),
    legend.position = "none",
    legend.text = element_text(size = 12, family = 'arial', face = 'bold'),
    legend.title = element_text(size = 12, family = 'arial', face = 'bold'),
    legend.position.inside = c(0.85, 0.5),
    aspect.ratio = 1,
    axis.ticks.length = unit(-1.25, "mm"),
    axis.ticks = element_line(size = 0.75),
    axis.line = element_line(size = 0.75, lineend = 'round'),
    legend.background = element_blank(),
    legend.box.background = element_rect(colour = "black", size = 0.75),
    legend.key.spacing.y = unit(1, 'mm'),
    legend.key.width = unit(2, 'mm'),
    axis.text.x = element_text(angle = 30, vjust = 0.8, hjust=0.91),
    panel.border = element_blank(),
  )
save_plot('hESC_APS_PXM_SMT_HAND1_log2_1p.png', p, base_width = 5, base_height = 5, dpi = 300)

p <- loh_RNA %>%
  filter(str_detect(sample, 'hESC|APS|DLL1PXM|D3Early')) %>% 
  mutate(sample = ordered(sample, levels = c('hESC', 'APS', 'DLL1PXM', 'D3EarlySmt'))) %>% 
  filter(gene == 'FOXC2', str_detect(batch, 'batch 2|batch 1')) %>% 
  ggplot(aes(x = sample, y = log21p_tpm)) +
  geom_star(starshape = 15, fill = '#00bbd7', color = '#077fb0', size = 3, starstroke = 1, alpha = 0.75) +
  geom_line(aes(group = gene, y = mean_log21p_tpm), color = '#00468BFF', alpha = 0.75, linewidth = 1) +
  labs(
    y=expression(bold(log["2"] * ("TPM + 1"))),
    title = 'FOXC2'
  ) +
  theme_pubr() +
  theme(
    text = element_text(size = 20, family = 'arial', face = 'bold'),
    legend.position = "none",
    legend.text = element_text(size = 12, family = 'arial', face = 'bold'),
    legend.title = element_text(size = 12, family = 'arial', face = 'bold'),
    legend.position.inside = c(0.85, 0.5),
    aspect.ratio = 1,
    axis.ticks.length = unit(-1.25, "mm"),
    axis.ticks = element_line(size = 0.75),
    axis.line = element_line(size = 0.75, lineend = 'round'),
    legend.background = element_blank(),
    legend.box.background = element_rect(colour = "black", size = 0.75),
    legend.key.spacing.y = unit(1, 'mm'),
    legend.key.width = unit(2, 'mm'),
    axis.text.x = element_text(angle = 30, vjust = 0.8, hjust=0.91),
    panel.border = element_blank(),
  )
save_plot('hESC_APS_PXM_SMT_FOXC2_log2_1p.png', p, base_width = 5, base_height = 5, dpi = 300)

p <- loh_RNA %>%
  filter(str_detect(sample, 'hESC|APS|DLL1PXM|D3Early')) %>% 
  mutate(sample = ordered(sample, levels = c('hESC', 'APS', 'DLL1PXM', 'D3EarlySmt'))) %>% 
  filter(gene == 'MSGN1', str_detect(batch, 'batch 2|batch 1')) %>% 
  ggplot(aes(x = sample, y = log21p_tpm)) +
  geom_star(starshape = 15, fill = '#00bbd7', color = '#077fb0', size = 3, starstroke = 1, alpha = 0.75) +
  geom_line(aes(group = gene, y = mean_log21p_tpm), color = '#00468BFF', alpha = 0.75, linewidth = 1) +
  labs(
    y=expression(bold(log["2"] * ("TPM + 1"))),
    title = 'MSGN1'
  ) +
  theme_pubr() +
  theme(
    text = element_text(size = 20, family = 'arial', face = 'bold'),
    legend.position = "none",
    legend.text = element_text(size = 12, family = 'arial', face = 'bold'),
    legend.title = element_text(size = 12, family = 'arial', face = 'bold'),
    legend.position.inside = c(0.85, 0.5),
    aspect.ratio = 1,
    axis.ticks.length = unit(-1.25, "mm"),
    axis.ticks = element_line(size = 0.75),
    axis.line = element_line(size = 0.75, lineend = 'round'),
    legend.background = element_blank(),
    legend.box.background = element_rect(colour = "black", size = 0.75),
    legend.key.spacing.y = unit(1, 'mm'),
    legend.key.width = unit(2, 'mm'),
    axis.text.x = element_text(angle = 30, vjust = 0.8, hjust=0.91),
    panel.border = element_blank(),
  )
save_plot('hESC_APS_PXM_SMT_MSGN1_log2_1p.png', p, base_width = 5, base_height = 5, dpi = 300)
```


```{r}
p <- loh_RNA %>%
  filter(str_detect(sample, 'hESC|APS|DLL1PXM')) %>% 
  mutate(sample = ordered(sample, levels = c('hESC', 'APS', 'DLL1PXM'))) %>% 
  ggplot(aes(x = sample, y = tpm)) +
  geom_star(
    aes(starshape = batch, fill = gene), size = 3, starstroke = 0.7, alpha = 0.75
  ) +
  geom_line(aes(group = interaction(batch, gene), color = gene), linetype = 'dashed') +
  theme_pubr() +
  scale_fill_npg() +
  scale_color_npg() +
  theme(
    text = element_text(size = 20, family = 'arial', face = 'bold'),
    legend.position = "inside",
    legend.text = element_text(size = 12, family = 'arial', face = 'bold'),
    legend.title = element_text(size = 12, family = 'arial', face = 'bold'),
    legend.position.inside = c(0.85, 0.5),
    aspect.ratio = 1,
    axis.ticks.length = unit(-1.25, "mm"),
    axis.ticks = element_line(size = 0.75),
    axis.line = element_line(size = 0.75, lineend = 'round'),
    legend.background = element_blank(),
    legend.box.background = element_rect(colour = "black", size = 0.75),
    legend.key.spacing.y = unit(1, 'mm'),
    legend.key.width = unit(2, 'mm'),
    panel.border = element_blank(),
  ) +
  labs(
    starshape = 'batch',
    color = 'gene',
    fill = 'gene'
  )
save_plot('hESC_APS_PXM_EOMES_FOXF1.png', p, base_width = 5, base_height = 5, dpi = 300)
```

```{r}
celltype <- df$Factor.Value.tissue.
celltype[nchar(celltype)==0] <- 'ESC'
modcombat <- model.matrix(~as.factor(celltype))
combatExpr <- ComBat(
  dat = logBulkTPM, 
  batch = df$Comment.batch., 
  mod = modcombat, 
  par.prior = FALSE, 
  prior.plots = FALSE)

combatExpr <- ComBat(
  dat = bulkTPMData, 
  batch = df$Comment.batch., 
  mod = modcombat, 
  par.prior = FALSE, 
  prior.plots = FALSE)

combatExpr[combatExpr < 0] <- 0
```


```{r}
cbind(bulkTPMData[, 1:2], combatExpr) %>% write.csv('combat_expression.csv', row.names = FALSE)
```

